Performance Models for Data Transfers: A Case Study with Molecular Chemistry Kernels

### 2019-04-09 ICPP-19/communications.tex
rearranging stuff after Lionel's modification

%%to comment all todos
%\renewcommand{\todo}{}

@@

\newcommand{\scomm}{\ensuremath{{S}_{\text{COMM}}}}
\newcommand{\scomp}{\ensuremath{{S}_{\text{COMP}}}}

@@

	%
	% The abstract is a short summary of the work to be presented in the article.
	\begin{abstract}
		With increasing complexity of hardwares, systems with different memory nodes are ubiquitous in high performance computing (HPC). It is paramount to develop strategies to overlap the data transfers between memory nodes with the computation in order to exploit the full potential of these systems. In this article, we consider the problem of deciding the  order of data transfers between two memory nodes for a set of independent tasks  with the objective to minimize the makespan. We propose an optimal heuristic to obtain the order of data transfers when there is not any restriction on memory limit. We also show that with finite memory capacity obtaining the optimal order of data transfers is an NP-complete problem. We propose several heuristics for this problem and provide details about their favorable situations. We present analysis of our heuristics on traces, obtained by running 2 molecular chemistry kernels, namely, Hartree–Fock (HF)  and Coupled Cluster Single Double (CCSD) on 10 nodes of an HPC system. Our results show that some of our heuristics achieve significant overlap for moderate memory capacities and are very close to the lower bound of makespan.
	\end{abstract}

	%%\begin{CCSXML}
	%%	<ccs2012>
	%%	<concept>
	%%	<concept_id>10010520.10010553.10010562</concept_id>
	%%	<concept_desc>Computer systems organization~Embedded systems</concept_desc>
	%%	<concept_significance>500</concept_significance>
	%%	</concept>
	%%	<concept>
	%%	<concept_id>10010520.10010575.10010755</concept_id>
	%%	<concept_desc>Computer systems organization~Redundancy</concept_desc>
	%%	<concept_significance>300</concept_significance>
	%%	</concept>
	%%	<concept>
	%%	<concept_id>10010520.10010553.10010554</concept_id>
	%%	<concept_desc>Computer systems organization~Robotics</concept_desc>
	%%	<concept_significance>100</concept_significance>
	%%	</concept>
	%%	<concept>
	%%	<concept_id>10003033.10003083.10003095</concept_id>
	%%	<concept_desc>Networks~Network reliability</concept_desc>
	%%	<concept_significance>100</concept_significance>
	%%	</concept>
	%%	</ccs2012>
	%%\end{CCSXML}
	\ccsdesc[100]{Computer systems organization~Embedded systems}
	\ccsdesc[100]{Computer systems organization~Redundancy}
	\ccsdesc{Computer systems organization~Robotics}
	\ccsdesc[100]{Networks~Network reliability}

@@

	Certain applications such as dense linear algebra kernels have regular structure. Therefore, it is possible to associate priorities, based on the task graph structure, with computations, which runtime uses to make the execution efficient. In irregular applications, programmers do not know the precise structure of the tasks graphs in advance, tasks are added recursively based on certain sentinel constraints. For such applications, runtime sees a set of independent ready tasks and schedule them on different processing units. It is extremely important for runtimes to decide the order of data transfers for these scheduled computations such that overlap between computations and communications can be maximised. This is the main topic of this article. We prove that order of communications on two memory nodes  with the objective of minimizing the makespan is  NP-Complete. Our proof is inspired from work by ~\cite{Papadimitriou:1980:FSL:322203.322213}, which applies a similar technique for 2-machine flowshop problem. The main difference with their and our approach is that they consider unit buffer requirement on 1st machine after job completion. On the contrary, our approach is designed where there is not any restriction on requested input data and memory is acquired before starting the data transfer on communication resource. We propose different strategies for runtimes such that it can maximize the overlap of computation and communication for kernels running on two different memory nodes. We also prove that one of our proposed strategies is an optimal one when we do not have any restriction on the capacity of memory node.  We evaluate our strategies on a cluster of homogeneous nodes. But our approach is generic and easily adaptable to any system which operates on different memory spaces. Here are the important contributions of this paper.

	\renewcommand{\labelitemi}{$\bullet$}
	\begin{itemize}
		\item Proposed different scheduling strategies with the objective to minimize the makespan
		\item An optimal strategy and its proof, when there is not any memory capacity constraint on the target machine
		\item NP-Completeness proof for the general data-transfer problem  
		\item Linear programming formulation of the problem
		\item Numerous  experiments to assess the effectiveness of our strategies on molecular chemistry kernels 
	\end{itemize}

	The outline of the article is the following. Section~\ref{sec:relatedWork} describes past work on the computations with limited memory and similar problems in the literature. In section~\ref{sec:theoreticalProof}, we present an algorithm to obtain the order of data transfers when there is not any memory capacity restriction and prove its optimality. Then, we also prove that in general data transfer problem is NP-complete. In Section~\ref{sec:heuristics}, we propose several heuristics and describe their favorable scenarios. We mainly consider three categories of heuristics: static heuristics, dynamic heuristics and static heuristics with dynamic corrections. We evaluate our proposed strategies on two molecular chemistry kernels in Section~\ref{sec:exp}.  Our results show that static heuristics with dynamic corrections achieve good performance in most cases. We finally  propose conclusions and perspectives in Section~\ref{sec:conclusion}.

@@

	The problem considered in this article also can be viewed as a flow shop problem. Communication and computation timings of a task can be considered as processing timings on different machines.  Johnson has provided scheduling strategies for 2 and 3-machine flow shop problems with infinite memory capacity~\cite{johnson}. 2-machine flow shop problem with finite memory capacity has been proven NP-Complete by Papadimitriou et. al~\cite{Papadimitriou:1980:FSL:322203.322213}. The main difference with the our approach is that they consider unit buffer requirement on 1st machine after job completion. On the contrary, our approach is designed for tasks appearing in scientific workloads whose sizes are highly irregular and memory is acquired before starting the communication.

@@

	To exploit the full potential of a system tasks may get scheduled on processing
	units where all of their data does not reside. A task may require all of its data
	in local memory before starting the execution. There may be multiple tasks
	scheduled on a processing unit, which require to transfer data from the same
	memory node. Order of data transfers for such tasks is very crucial for the
	communication-computation overlap, thus for the overall performance. Tasks may
	also produce some output data, which, in general, is very less.  It is often the case that future tasks running on the same memory node require output data of the past tasks. Therefore, most runtime systems transfer data to other memory nodes based on the demand -- not immediately. It is possible that all output data can be stored in a preallocated separate buffer on a memory node. Hence, we do not consider output data separately in our analysis and assume that output data is negligible  or stored in a separate buffer for each task. We prove that order of execution of such tasks with finite memory capacity is an NP-complete problem. We also present an optimal algorithm to determine the order of execution of tasks when there is not any  restriction on memory limit.

	\subsection{Infinite Memory Case}
	Let $S$ be the set of tasks assigned on a processor with memory unit $M$, and this set of tasks  need input data from another memory unit $M'$. We assume that size of $M$ is very large and there will not be any memory capacity restriction for tasks.

	Let \scomm($i$) and  \scomp($i$) represent the start times of task $i$ on communication and computation resources. $CM_i$ and $CP_i$ are communication and computation times of task $i$. We call a task $i$ is compute (resp. communication) intensive if $CP_i $$\ge$ (resp. $<$) $CM_i$ ).

	\begin{algorithm}
		\caption{\label{alg:OrderOfExecutionInfinteMemory}Algorithm to determine the order of processing and schedule for a set of ready tasks (infinite memory case).}
		\begin{algorithmic}[1]
			\STATE Divide ready tasks in two sets $S_1$ and $S_2$. If computation time of a task $T$is not less than its communication time, then $T$ is in $S_1$ otherwise in $S_2$.
			\STATE Sort $S_1$ in queue $Q$ by non-decreasing communication times
			\STATE Sort $S_2$ in queue $Q'$ by non-increasing computation times
			\STATE Append $Q'$ to $Q$
			\STATE $early\_available\_time\_on\_communication\_resource \gets 0$
			\STATE $early\_available\_time\_on\_computation\_resource \gets 0$
			\WHILE{$Q \neq \emptyset$}
			\STATE remove a task $T$ from beginning of $Q$ for processing
			\STATE $\scomm(T) \gets  early\_available\_time\_on\_communication\_resource$
			\STATE $\scomp(T) \gets  max(\scomm(T) + CM_T, early\_available\_time\_on\_computation\_resource)$
			\STATE $early\_available\_time\_on\_communication\_resource \gets  \scomm(T) + CM_T$
			\STATE $early\_available\_time\_on\_computation\_resource \gets  \scomp(T) + CP_T$
			\ENDWHILE
		\end{algorithmic}
	\end{algorithm}

\noindent In order to prove the optimality of makespan for the schedule obtained by the  Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}, we need the following lemma.

	\begin{lemma}\label{lemma:swappingOfTasks}
		Swapping two contiguous tasks $A$ and $B$ in a schedule does not improve the makespan if one of the conditions is true.
		\begin{enumerate}[label=\roman*)]
			\item  $CP_A \ge CM_A, CP_B \ge CM_B, CM_A \le CM_B$
			\item $CP_A < CM_A, CP_B < CM_B, CP_A \ge CP_B$
			\item $CP_A \ge CM_A, CP_B < CM_B$
		\end{enumerate}
	\end{lemma}
	\begin{proof}
		Let $t_1$ and $t_2$ be the early start time on communication and computation resources just before the task $A$ starts.

		\begin{figure}[htb]
			\centering
			\includegraphics[scale=0.35]{Figs/original_swapped_schedules}
			\caption{ \label{fig:bothSchedule} Original and Swapped Schedules.}
		\end{figure}
		\todo[inline]{SKumar: decrease space between both schedules of Figure~\ref{fig:bothSchedule}}
		We write the following constraints based on the two schedules of the Fig~\ref{fig:bothSchedule}.

			\noindent $\scomp(A) = max(t_1 + CM_A, t_2)$ \\
			\noindent $\scomp(B) = max(\scomp(A) + CP_A,  t_1 + CM_A + CM_B)$\\

%%		\begin{eqnarray*}
%%			\scomp(A) & = & max(t_1 + CM_A, t_2) \\
%%			\scomp(B) & = & max(\scomp(A) + CP_A,  t_1 + CM_A + CM_B)
%%		\end{eqnarray*}

		\noindent Completion time of $B$ in original Schedule = $\scomm(B) + CP_B$\\
		\noindent Completion time of $B$ in swapped Schedule,  $\scomp '(B)  =  max(t_1 + CM_B, t_2)$\\
		\noindent Completion time of $A$ in swapped Schedule, $ \scomp '(A) = max(\scomp'(B) + CP_B, t_1 + CM_B + CM_A)$

%%		\begin{eqnarray*}
%%			\text{Completion time of $B$ in swapped Schedule, } \scomp '(B) & = & max(t_1 + CM_B, t_2)\\
%%			\text{Completion time of $A$ in swapped Schedule, } \scomp '(A) & = & max(\scomp'(B) + CP_B, t_1 + CM_B + CM_A)
%%		\end{eqnarray*}

		In both schedules, early available time on communication resource after scheduling $A$ and $B$ is same. If we show that early available time on computation resource in swapped schedule after scheduling both tasks is not less than the time of original schedule, then we are done. Hence our goal is to prove that,		
		$$\scomp(B) + CP_B \le \scomp'(A) + CP_A $$

		\begin{eqnarray*}
			\scomp(B) + CP_B & = & max(\scomp(A) + CP_A + CP_B,  t_1 + CM_A + CM_B + CP_B)\\
			& = & max(t_1 + CM_A + CP_A + CP_B, t_2 + CP_A + CP_B, t_1 + CM_A + CM_B + CP_B) 
		\end{eqnarray*}
		Case I: $CP_A \ge CM_A, CP_B \ge CM_B, CM_A \le CM_B$
		\begin{eqnarray*}
			\scomp(B) + CP_B & = & max(t_1 + CM_A + CP_A + CP_B, t_2 + CP_A + CP_B, t_1 + CM_A + CM_B + CP_B)\\
			& \le& max(t_1 + CM_A + CP_B, t_2  + CP_B, t_1  + CM_B + CP_B) + CP_A\\
			& = & max( t_1  + CM_B + CP_B, t_2  + CP_B) + CP_A \\
			& = & max(max(t_1 + CM_B, t_2) + CP_B, t_1 + CM_A + CM_B) + CP_A\\
			& = & max(\scomp'(B) + CP_B, t_1 + CM_A + CM_B) + CP_A \\
			& = & \scomp'(A) + CP_A
		\end{eqnarray*}
		Case II: $CP_A < CM_A, CP_B < CM_B, CP_A \ge CP_B$
		\begin{eqnarray*}
			\scomp(B) + CP_B & = & max(t_1 + CM_A + CP_A + CP_B, t_2 + CP_A + CP_B, t_1 + CM_A + CM_B + CP_B)\\
			& \le & max(t_1 + CM_A + CP_B, t_2+CP_B, t_1 + CM_A + CM_B) + CP_A\\
			& = & max(t_1 + CM_A + CM_B, t_2+ CP_B) + CP_A \\
			& = & max(t_1 + CM_B + CP_B, t_2 + CP_B, t_1 + CM(A) + CM_B) + CP_A \\
			& = & max(max(t_1 + CM_B, t_2) + CP_B, t_1 + CM_A + CM_B) + CP_A \\
			& = & max(\scomp'(B) + CP_B, t_1 + CM_A +CM_B) + CP_A \\
			& = & \scomp'(A) + CP_A
		\end{eqnarray*}
		Case III: $CP_A \ge CM_A, CP_B < CM_B$
		\begin{eqnarray*}
			\scomp(B) + CP_B & = & max(t_1 + CM_A + CP_A + CP_B, t_2 + CP_A + CP_B, t_1 + CM_A + CM_B + CP_B)\\
			& \le & max(t_1 + CM_A + CP_B, t_2 + CP_B, t_1 + CM_B + CP_B) + CP_A\\
			& \le & max(t_1 + CM_A + CM_B, t_2 + CP_B, t_1 + CM_B + CP_B) + CP_A \\
			& = & max(max(t_1 + CM_B, t_2) + CP_B, t_1 + CM_A + CM_B) + CP_A \\
			& = & max(\scomp'(B) + CP_B, t_1 + CM_A +CM_B) + CP_A \\
			& = & \scomp'(A) + CP_A
		\end{eqnarray*}
	\end{proof}

	\begin{theorem}
		Scheduled constructed by Algorithm~\ref{alg:OrderOfExecutionInfinteMemory} achieves optimal makespan.
	\end{theorem}
	%%	\begin{proof}
	%%		Let us first define some notations which we need for this proof.
	%%%%		\todo[inline]{SKumar: Specify the destination for all notations}
	%%		\begin{align*}
	%%		t : & \text{Time at which data transfer for task A starts in original schedule}\\
	%%		S_A : & \text{ Start of computation  time of task A in original schedule}\\
	%%		S_B : & \text{ Start of computation time of task B in original schedule}\\
	%%		S'_A : & \text{ Start of computation time of task A in swapped schedule}\\
	%%		S'_B : & \text{ Start of computation time of task B in swapped schedule}\\
	%%		CM(A) : & \text{ Data transfer time for task A}\\
	%%		CM(B) : & \text{ Data transfer time for task B}\\
	%%		CP(A) : & \text{ Computation time of of task A}\\
	%%		CP(B) : & \text{ Computation time of task B}\\
	%%		\rho : & \text{ ratio of length of task and data transfer time}
	%%		\end{align*}
	%%	\end{proof}

	\begin{proof}

		Let $O$ be an optimal schedule. We assume that $O$ is a permuation schedule, If it is not the case, we make the order of computations same as the order of communications without increasing the makespan.  Suppose two tasks have opposite order on both resources then we position the second task just before the first task on the computation resource. It is evident that this change does not alter the optimal  makespan, we repeat this procedure until order of communications and computations is same in $O$.

		Let $S$ be the  schedule obtained from Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}. We prove the theorem by converting $S$ to $O$ and showing that at each step makespan of intermediate schedule is not less than the original makespan. We rely on Lemma~\ref{lemma:swappingOfTasks} to convert $S$ to $O$.

		We traverse schedule $O$ from left to right and for each $ith$ task in sequence, we apply Lemma~\ref{lemma:swappingOfTasks} repetitively until order of task in the swapped schedule is same. It is obvious that after moving the $ith$ task at the beginning remaining schedule (schedule after $ith$ task) still satisfies one of the conditions of Lemma~\ref{lemma:swappingOfTasks}.

		Let the final swapped schedule is $S_{final}$. From Lemma~\ref{lemma:swappingOfTasks}, $makespan(S_{final})$ $ \ge $ $makesapn(S)$. From the construction, $makespan(S_{final})$ $ \le$ $ makesapn(O)$. As $O$ is an optimal schedule, hence $makespan(S) = makesapn(O)$. This completes our proof.
	\end{proof}

	%%	 
	%%	Let $S$ be the schedule obtained by the above heuristic and $O$ be an optimal schedule. We prove the above claim by converting $S$ to $O$ and showing that at each step completing time of intermediate schedule is not less than the completion time of original schedule $S$. In order to do so, we make sure order of two tasks in both schedules $S$ and $O$ is same. If not, We perform swapping of two tasks.
	%%	
	%%	\noindent If we prove the swapping strategy for two contiguous tasks holds, then the same strategy transitively hold for any two tasks of Schedule $S$

>>>>

%%to comment all todos
%%\renewcommand{\todo}{}

@@

\newcommand{\scomm}{\ensuremath{{S}_{\text{COMM}}}}
\newcommand{\scomp}{\ensuremath{{S}_{\text{COMP}}}}
\newcommand{\threepart}{\textsc{3Par}\xspace}

@@

	%
	% The abstract is a short summary of the work to be presented in the article.
	\begin{abstract}
		With increasing complexity of hardwares, systems with different memory nodes are ubiquitous in high performance computing (HPC). It is paramount to develop strategies to overlap the data transfers between memory nodes with computations in order to exploit the full potential of these systems. In this article, we consider the problem of deciding the order of data transfers between two memory nodes for a set of independent tasks with the objective to minimize the makespan. We prove that with finite memory capacity obtaining the optimal order of data transfers is an NP-complete problem. We propose several heuristics for this problem and provide details about their favorable situations. We present analysis of our heuristics on traces, obtained by running 2 molecular chemistry kernels, namely, Hartree–Fock (HF) and Coupled Cluster Single Double (CCSD) on 10 nodes of an HPC system. Our results show that some of our heuristics achieve significant overlap for moderate memory capacities and are very close to the lower bound of makespan.
	\end{abstract}

	%%\begin{CCSXML}
	%%	<ccs2012>
	%%	<concept>
	%%	<concept_id>10010520.10010553.10010562</concept_id>
	%%	<concept_desc>Computer systems organization~Embedded systems</concept_desc>
	%%	<concept_significance>500</concept_significance>
	%%	</concept>
	%%	<concept>
	%%	<concept_id>10010520.10010575.10010755</concept_id>
	%%	<concept_desc>Computer systems organization~Redundancy</concept_desc>
	%%	<concept_significance>300</concept_significance>
	%%	</concept>
	%%	<concept>
	%%	<concept_id>10010520.10010553.10010554</concept_id>
	%%	<concept_desc>Computer systems organization~Robotics</concept_desc>
	%%	<concept_significance>100</concept_significance>
	%%	</concept>
	%%	<concept>
	%%	<concept_id>10003033.10003083.10003095</concept_id>
	%%	<concept_desc>Networks~Network reliability</concept_desc>
	%%	<concept_significance>100</concept_significance>
	%%	</concept>
	%%	</ccs2012>
	%%\end{CCSXML}
	\ccsdesc[100]{Computer systems organization~ High Performance Computing }
	\ccsdesc[100]{Computing Methodologies~Modeling and Simulation}
	\ccsdesc[100]{General ~ Performance}

@@

	Certain applications such as dense linear algebra kernels have regular structure. Therefore, it is possible to associate priorities, based on the task graph structure, with computations, which runtime uses to make the execution efficient. In irregular applications, programmers do not know the precise structure of the task graphs in advance, tasks are added recursively based on certain sentinel constraints. For such applications, runtime sees a set of independent ready tasks and schedule them on different processing units. It is extremely important for runtimes to decide the order of data transfers for these scheduled computations such that overlap between computations and communications can be maximised. This is the main topic of this article. We prove that order of communications on two memory nodes with the objective of minimizing the makespan is NP-Complete. Our proof is inspired from work by ~\cite{Papadimitriou:1980:FSL:322203.322213}, which applies a similar technique for 2-machine flowshop problem. The main difference with their and our approach is that they consider a bounded number of tasks can await execution on second machine. On the contrary, our approach is designed for tasks appearing in scientific workloads whose memory requirements are highly irregular and memory is acquired before starting the data transfer on communication resource. We propose different strategies for runtimes such that it can maximize the overlap of computation and communication for kernels running on two different memory nodes. We evaluate our strategies on a cluster of homogeneous nodes. But our approach is generic and easily adaptable to any system which operates on different memory spaces. Here are the important contributions of this paper.

	\renewcommand{\labelitemi}{$\bullet$}
	\begin{itemize}
		\item Proposed different scheduling strategies with the objective to minimize the makespan
%%		\item An optimal strategy and its proof, when there is not any memory capacity constraint on the target machine
		\item NP-Completeness proof for the general data-transfer problem 
		\item Linear programming formulation of the problem
		\item Numerous experiments to assess the effectiveness of our strategies on molecular chemistry kernels 
	\end{itemize}

	The outline of the article is the following. Section~\ref{sec:relatedWork} describes past work on the computations with limited memory and similar problems in the literature. In section~\ref{sec:theoreticalProof}, we present an algorithm to obtain the order of data transfers when there is not any memory capacity restriction. Then, we also prove that in general data transfer problem is NP-complete. In Section~\ref{sec:heuristics}, we propose several heuristics and describe their favorable scenarios. We mainly consider three categories of heuristics: static heuristics, dynamic heuristics and static heuristics with dynamic corrections. We evaluate our proposed strategies on two molecular chemistry kernels in Section~\ref{sec:exp}. Our results show that static heuristics with dynamic corrections achieve good performance in most cases. We finally propose conclusions and perspectives in Section~\ref{sec:conclusion}.

@@

	The problem considered in this article also can be viewed as a flow shop problem. Communication and computation timings of a task can be considered as processing timings on different machines. Johnson has provided scheduling strategies for 2 and 3-machine flow shop problems with infinite memory capacity~\cite{johnson}. 2-machine flow shop problem with finite buffer has been proven NP-Complete by Papadimitriou et. al~\cite{Papadimitriou:1980:FSL:322203.322213}. They consider a bounded number of tasks can await execution on second machine.

@@

	To exploit the full potential of a system tasks may get scheduled on processing
	units where all of their data does not reside. A task may require all of its data
	in local memory before starting the execution. There may be multiple tasks
	scheduled on a processing unit, which require to transfer data from the same
	memory node. Order of data transfers for such tasks is very crucial for the
	communication-computation overlap, thus for the overall performance. In general, order of task execution with input and output data transfers can be 
	viewed as 3-machine flowshop problem, where processing time on first machine is 
	input data transfer time, processing time on second machine is task execution time, 
	processing time on third machine is output data transfer time and the objective is 
	to minimize the total makespan. This is a well known hard problem~\cite{NPComplete3Machine,johnson}.

	In general, output data produced by a task is very less. It is often the case that future tasks running on the same memory node require output data of the past tasks. Therefore, most runtime systems transfer data to other memory nodes based on the demand -- not immediately. It is possible that all output data can be stored in a preallocated separate buffer on a memory node. Hence, we do not consider output data separately in our analysis and assume that output data is negligible or stored in a separate buffer for each task. Thus problem considered here is simpler than general 3-machine flowshop problem. We prove that order of execution of such tasks with finite memory capacity is an NP-complete problem. We also present an optimal algorithm to determine the order of execution of tasks when there is not any restriction on memory limit.

	\todo[inline]{Find a citation saying 3-machine flowshop problem is NP-Complete.}
%%%%	done:\todo[inline]{Rewrite both paragraphs together.}
%%	In general, order of task execution with input and output data transfers can be 
%%	viewed as 3-machine flowshop problem, where processing time on 1st machine is 
%%	input data transfer time, processing time on 2nd machine is task execution time, 
%%	processing time on 3rd machine is output data transfer time and the objective is 
%%	to minimize the total makespan. This is a well known hard problem~\cite{NPComplete3Machine,johnson}. As stated earlier, we do not consider output data separately in our analysis and assume that output data is negligible or stored in a separate buffer. Hence, problem considered in this article is simpler than general 3-machine flowshop problem.

	\noindent\textbf{Problem $DT$} : A set of tasks $ST=\{T_1,
	\cdots, T_n\}$ is scheduled on a processing unit $P$ with
	memory unit $M$ of capacity $C$. Input data for tasks of $ST$
	reside on another memory unit $M'$. $COMM_i$ is the time to
	transfer input data from $M'$ to $M$ for task $i$ and $COMP_i$
	is the execution time of task $i$ on $P$. We assume that these
	tasks do not produce any output data. There can be only one
	communication at a time, and $P$ can only process one task at
	a time. A task uses an amount of memory in memory $M$ from the
	start of its communication to the end of its computation.

	\noindent Given $L$, is there a feasible schedule $S$ for $ST$ such that
	makespan of $S$, $\mu(S) \le L$?

%%	\todo[inline]{Mention that memory occupation of a task is the same as its communication time. Skumar: it is not required at all, but we use this for simplicity in Theorem~\ref{th:npComplete}}
%%	\todo[inline]{Mention that it is always best to use the same order on both machines.}

Let \scomm($i$) and  \scomp($i$) represent the start times of task $i$ on communication and computation resources. We call a task $i$ is compute (resp. communication) intensive if $CP_i $$\ge$ (resp. $<$) $CM_i$ ).

	\subsection{Special case: Infinite Memory}

	When the computing resource has a very large memory, our problem
	becomes a classic 2-machine flowshop problem: communication time is
	the processing time on the first machine and computation time is the
	processing time on the second machine. Johnson's
	algorithm~\cite{johnson} is known to provide an ordering for the tasks
	which results in an optimal makespan. This algorithm is rewritten in
	Algorithm~\ref{alg:OrderOfExecutionInfinteMemory}.

	\begin{algorithm}
		\caption{\label{alg:OrderOfExecutionInfinteMemory}Johnson's~\cite{johnson} algorithm (infinite memory case).}
		\begin{algorithmic}[1]
			\STATE Divide ready tasks in two sets $S_1$ and $S_2$. If computation time of a task $T$ is not less than its communication time, then $T$ is in $S_1$ otherwise in $S_2$.
			\STATE Sort $S_1$ in queue $Q$ by non-decreasing communication times
			\STATE Sort $S_2$ in queue $Q'$ by non-increasing computation times
			\STATE Append $Q'$ to $Q$
			\STATE $\tau_{\text{COMM}} \gets 0$ \hfill\COMMENT{Available time of communication resource}
			\STATE $\tau_{\text{COMP}} \gets 0$\hfill \COMMENT{Available time of computation resource}
			\WHILE{$Q \neq \emptyset$}
			\STATE Remove a task $T$ from beginning of $Q$ for processing
			\STATE $\scomm(T) \gets \tau_{\text{COMM}}$
			\STATE $\scomp(T) \gets max(\scomm(T) + CM_T, \tau_{\text{COMP}})$
			\STATE $\tau_{\text{COMM}} \gets \scomm(T) + CM_T$
			\STATE $\tau_{\text{COMP}} \gets \scomp(T) + CP_T$
			\ENDWHILE
		\end{algorithmic}
	\end{algorithm}

	We also prove optimality of Algorithm~\ref{alg:OrderOfExecutionInfinteMemory} in a different way, which is available in an extended version(cite).
	\subsection{Finite Memory}

	We now consider the general case, in which the memory limit is a
	constraint for the schedule. This is related to previous work by
	Papadimitriou et. al~\cite{Papadimitriou:1980:FSL:322203.322213}, in which the second machine
	can only handle a bounded number of tasks. Our problem generalizes this
	work to heterogeneous memory consumption among tasks, with an
	additional difference: memory usage starts at the beginning of the
	first part of a task (instead of at the end of the first part). This
	requires to provide a slightly different NP-completeness proof, as
	given below.

	\begin{theorem}\label{th:npComplete}
		Problem $DT$ is NP-complete.
	\end{theorem}
	\begin{proof}

		It is easy to see that the $DT$ belongs in NP: given a schedule, one
		can check in linear time that at each start of a communication, the
		memory constraint is satisfied, and that task starts execution only
		after its input data is transferred to $M$.

		In order to prove NP-hardness, we use a reduction from the
		well-known NP-complete problem 3 Partition~\cite{GareyJohnson}:

		\noindent \textbf{Three Partition Problem} (\threepart): Given a set of
		$3m$ integers $A = \{ a_1, \cdots, a_{3m }\}$ with $a_i > 1$, is there a partition
		of $A$ into $m$ triplets $TR_i = \{a_{i_1}, a_{i_2}, a_{i_3}\}$,
		such that $\forall i, a_{i_1} + a_{i_2} + a_{i_3} = b$, where
		$b=(1/m) \sum a_i $?

		Let us first show that \threepart problem reduces in polynomial time to
		problem $DT$. Suppose that we are given an instance $A = \{ a_1,
		\cdots, a_{3m }\}$ of \threepart. It is immediately obvious that
		$a_i>1$, since we can always add sufficiently large integers to
		the $a_i$ values and scale the problem accordingly. This scaling will not
		affect in any way the existence of a solution for the instance of
		\threepart problem.

		From such an instance, we define $x = max\{a_i:1\le i\le 3m\}$,
		and we construct an instance $I$ of the problem $DT$ with $4m+1$
		tasks, whose characteristics are given in
		Table~\ref{table:np.completeness.tasks}.

		\begin{table}[htb]
			\begin{tabular}{ |c|c|c| }
				\hline
				Task & Transfer time ($comm_i$)& Computation time ($e_i$)\\ \hline
				$K_0$ & $0$ & $3$ \\ \hline
				$K_1, \cdots, K_{m-1}$ & $b'=b+6x$ & $3$\\ \hline
				$K_m$ & $b'=b+6x$ & $0$ \\ \hline
				$1\le i \le 3m, A_i$ & $1$ & $a_i' = a_i + 2x$\\ \hline
			\end{tabular}
			\noindent Memory capacity: $C=b'+3$\\
			\noindent Target makespan: $L=m(b'+3)$
			\caption{Definition of tasks in the reduction from \threepart.}
			\label{table:np.completeness.tasks}
		\end{table}

		$Transfer time = Latency + \frac{Data Volume}{Bandwidth}$. For simplicity we consider $Latency=0$ and $Bandwidth=1$, (implies Input data volume = Input transfer time). Otherwise, we can adjust $C$ in a way such that at any point in a schedule at max one $K_i$ and three $A_i$ tasks can be active.

		We show that $I$ has a schedule $S$ with makespan at most $L$ if and
		only if the original \threepart instance has a solution. Notice that
		the sum of transfer times and the sum of computation times are both
		equal to $L$, therefore a valid schedule of makespan at most $L$ has
		makespan exactly $L$, with no idle time on both resources. It
		indicates that the first task is $K_0$ and the last task is $K_m$.

		\begin{figure}[htb]
			\tikzset{xtick/.style={inner xsep=0pt, inner ysep=3pt, minimum size=0pt, draw},%
				task/.style args={#1start#2length#3res#4color#5}{rounded corners, draw, inner
					sep=0pt, fill=#5, label=center:#1, fit={(#2,#4*0.75) (#2+#3,#4*0.75+0.75)}},%
				vert/.style={inner sep=1pt, fill=black, circle, draw, label=#1}
			}
			\newcommand{\schedule}[1]{
				\draw[->] (-0.4, 0) -- (#1, 0) node[below] {$t$};
				\draw (0, 0) -- (0, 1.5) node[pos=0.25, left] {Comp.}
				node[pos=0.75, left] {Comm.};
				\draw[dashed,gray] (0, 0.75) -- (#1, 0.75);
			}
			\centering
			\begin{tikzpicture}[yscale=0.7, thick, xscale=0.6]
			\schedule{12.5}
			\node[task=$A_{1,1}$ start 0 length 1 res 1 color cyan]{};
			\node[task=$A_{1,2}$ start 1 length 1 res 1 color blue!40!white]{};
			\node[task=$A_{1,3}$ start 2 length 1 res 1 color blue!70!white]{};
			\node[task=$K_0$ start 0 length 3 res 0 color gray!40!white]{};
			\node[task=$K_1$ start 3 length 6 res 1 color green]{}; 
			\node[task=$A_{1,1}$ start 3 length 1.8 res 0 color cyan]{};
			\node[task=$A_{1,2}$ start 4.8 length 2.3 res 0 color blue!40!white]{};
			\node[task=$A_{1,3}$ start 7.1 length 1.9 res 0 color blue!70!white]{};
			\node[task=$A_{2,1}$ start 9 length 1 res 1 color cyan]{};
			\node[task=$A_{2,2}$ start 10 length 1 res 1 color blue!40!white]{};
			\node[task=$A_{2,3}$ start 11 length 1 res 1 color blue!70!white]{};
			\node[task=$K_1$ start 9 length 3 res 0 color green]{};
			\draw[<->,thin] (0, -0.2) -- node[below]{$3$} (3, -0.2) ;
			\draw[<->,thin] (9, -0.2) -- node[below]{$3$} (12, -0.2) ;
			\draw[<->,thin] (3, -0.2) -- node[below]{$b'$} (9, -0.2) ;
			\end{tikzpicture}
			\caption{ \label{fig:firstSegment} Pattern of feasible schedule $S$.}
		\end{figure}

		If the \threepart instance has a solution, $A$ can be partitioned into
		$m$ triplets $TR_i = \{a_{i_1}, a_{i_2}, a_{i_3}\}$ such that $\forall
		i, a_{i_1} + a_{i_2} + a_{i_3} = b$, then we can construct a feasible
		schedule $S$ without idle times by the pattern depicted in
		Figure~\ref{fig:firstSegment}. The communications of tasks in $TR_i$ take
		place during the computation of task $K_{i-1}$, and the computations
		of tasks in $TR_i$ take place during the communication of task
		$K_i$. Since the memory capacity is $C=b'+3$, all tasks from a triplet
		can fit in memory with a task $K_i$, and their duration are exactly
		equal to the communication time of $K_i$. This schedule is thus
		feasible, and has length exactly $L$.

		\medskip

		We now prove that any feasible schedule of $I$ corresponds to a valid
		decomposition of $A$ for \threepart. Indeed, we argue that every
		feasible schedule has to consist of $m$ segments like the one shown in
		Figure~\ref{fig:firstSegment}. Each segment provides a triplet
		$\{a_{i_1}, a_{i_2}, a_{i_3}\}$ such that $a_{i_1} + a_{i_2} + a_{i_3}
		= b$.

		Any schedule $S$ of $I$ having no idle time must start with $K_0$. We
		first show that no other $K_i$ task can be active with $K_0$,
		otherwise we would get idle time on computation channel. Indeed, the
		communication of such a task $K_i$ would end at time at least $b'>3 +
		6x$, but at most two $A_i$ tasks can be computed, and they end at time
		at most $3+2max\{a_i':1\le i\le 3m\} = 3 + 6x$

		Hence three $A_i$ tasks must follow $K_0$. The memory requirement of
		other $K_i$ tasks is $b'$ and $2b'>C$, therefore at any point in the
		schedule at most one $K_i$ task can be active. Since the total
		duration of all $K_i$ tasks is $3 + (m-1)(b'+3) + b' = m (b'+3)=L$, at
		each point in $S$ exactly one $K_i$ task is active.

		With these $K_i$ tasks in place, the schedule on the computation
		resource contains $m$ slots of length exactly $b'$, in which all $A_i$
		tasks must fit without preemption. We can thus define triplet $TR_i$
		as the setof tasks which execute during the communication phase of
		task $K_i$. Since at each point in $S$, exactly one $K_i$ task is
		active, and since $S$ has no idle time on the computation resource,
		the total execution time of tasks in $TR_i$ is exactly $b'$, and thus
		$a_{i_1} + a_{i_2} + a_{i_3} = b$. This partition is thus a valid
		solution for the \threepart instance $A$.

		%% We have shown that $3PAR$ problem reduces
		%% to $DT$, and hence problem $DT$ is NP-complete.
	\end{proof}
